---
title: "Getting to know my audience"
author: "An example of literate programming"
date: "October 7, 2016"
output: 
  html_document: 
    fig_caption: yes
    highlight: pygments
    theme: cerulean
---


```{r, warning=FALSE, message=FALSE}
# Load libraries
library(googlesheets)
library(dplyr);library(ggplot2)
library(knitr);library(xtable)
opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE, fig.align='center') 

```

# Introduction
This report analyzes [survey data](https://goo.gl/forms/y7V2IvC60fEujjRf1) collected 
at the Emphasizing Reproducible Research by Teaching Data Analysis with R Presentation 
at the CELT 2016 Conference. After setting up proper authorization tokens, we connect 
directly to Google Sheets to download the data. 
```{r}
gs_auth(token="googlesheets_token.rds")
sheet_info <- gs_title("CELT_Responses")
raw <- sheet_info %>% gs_read()
```

Looking at the raw data, we see that the names of the data frame are set
to the question titles, which makes analysis very difficult to deal with. 
We rename these variable names something easier to manage. 
```{r}
names(raw) <- c("timestamp", "college", "years_teaching", "years_chico", 
                "R_research", "R_teach", "IWantR")
```

# Description of the data set

## College

```{r}
tc <- table(raw$college)
mx <- which(tc==max(tc))
tp <- paste0(tc, " (", round(prop.table(tc)*100,1), ")%")
kable(t(tc))
```

```{r college, fig.height=4, fig.width=8, fig.cap="College Representation"}
ggplot(raw, aes(x=college)) + geom_bar()
```

Figure 1 shows that the majority of the audience (n=`r tp[mx]`) is from the college of `r names(tc)[mx]`.

## Usage of R
```{r}
rr <- table(raw$R_teach, raw$R_research)
rr.p <- round(prop.table(rr)*100,1)
rr.r <- round(prop.table(rr, margin=1)*100,1)
rr.c <- round(prop.table(rr, margin=2)*100,1)
rr.a <- addmargins(rr)
kable(rr.a)
```

There are `r NROW(raw)-rr[1,1]` (`r 100-rr.p[1,1]`%) people in the audience who use 
R for their teaching or research. Of the `r rr.a[3,2]` who use it for research, `r rr[1,2]` (`r rr.c[1,2]`%) 
don't use it for teaching. 

```{r}
ggplot(raw, aes(x=R_teach, fill=R_research)) + geom_bar(position=position_dodge())
```

## Years teaching
The correlation between years teaching overall and years living in Chico is `r round(cor(raw$years_chico, raw$years_teaching),2)`. The plot below demonstrates that this relationship is _positive_. 

```{r}
ggplot(raw, aes(x=years_chico, y=years_teaching)) + geom_point() + 
  geom_smooth() + xlab("Years living in Chico") + ylab("Years teaching")
```

Does this relationship differ by college? 
```{r}
ggplot(raw, aes(x=years_chico, y=years_teaching, col=college)) + geom_point() + 
  geom_smooth(se=FALSE, method='lm') + xlab("Years living in Chico") + ylab("Years teaching")
```

Now the big question: Is the use of R in research or teaching a relative new thing? 
```{r}
ggplot(raw, aes(y=years_teaching, x=R_teach)) + geom_boxplot() + 
  facet_wrap(~R_research) + 
  ggtitle("Distribution of years teaching grouped by whether or not R is taught, 
          stratified by if R is used in research")
```

An ANOVA was used to test if there is a significant difference in the average
years teaching for those who teach, or do research in R. 
```{r}
mod1 <- anova(lm(years_teaching ~ R_teach + R_research, data=raw))
pv <- c(unlist(mod1)[13:14])
sw <- ifelse(pv < 0.05, "significant", "non-significant")
```

```{r}
kable(mod1, digits=2)
```

The results show that there is a `r sw[1]` difference in the average age
of those who teach R and those who do not, and a `r sw[2]` difference
between those who do research with R and those who do not. 


